---
title: "Four Parameters Categorical"
author: Naveed Ahmad
execute: 
  echo: false
---

```{r}
#| message: false
library(tidyverse)
library(primer.data)
library(tidymodels)
library(broom)
library(dplyr)
library(gt)
library(marginaleffects)
library(ggplot2)
library(dplyr)
```

<div style="text-align: justify;">
## Summary 
Different groups of Americans participate in voting for Presidential elections. What was the difference in voting preference of male and female groups for voting in Presidential election of 1992 for three contesting candidates Bill Clinton, George H. W. Bush, and Ross Perot? We will answer this question using data from National Election Survey. However, our findings are based upon some assumptions that survey data is valid, stable, responded by all sections of society eligible for voting without any bias. The statistical model I have used is multinominal regression because regressor variable was having more than one outcome with sex as covariate. Regression coefficients were measured and on the basis of estimates of parameters, Predictions were found and result in the form of a chart shows that 52% female voters will prefer voting for Clinton with uncertainty that approximately voting preference of female group may decrease by 2% or increase by 6% in favor of Clinton.
</div>

```{r}
#| cache: true
nes_92 <- nes |> 
  filter(year == 1992) |> 
  select(sex, pres_vote) |> 
  drop_na() |> 
  mutate(pres_vote = as.factor(case_when(
    pres_vote == "Democrat" ~ "Clinton",
    pres_vote == "Republican" ~ "Bush",
    pres_vote == "Third Party" ~ "Perot",
  ))) 

fit_nes <- multinom_reg(engine = "nnet") |>
  fit(pres_vote ~ sex, data = nes_92)
```

## Regression Coefficients
```{r}
model_results <- tibble::tibble(
  y.level = c("Clinton", "Clinton", "Perot", "Perot"),
  term = c("(Intercept)", "sexMale", "(Intercept)", "sexMale"),
  estimate = c(0.4553900, -0.2554092, -0.8518724, 0.4202485),
  conf.low = c(0.3090152, -0.4726795, -1.0612590, 0.1376819),
  conf.high = c(0.60176476, -0.03813882, -0.64248586, 0.70281513))

# Round to 2 decimal places and create a gt table
model_results |>
  mutate(across(estimate:conf.high, ~ round(.x, 2))) |>
  gt() |>
  tab_header(
    title = "Fitted Multinomial Logistic Regression Coefficients") |>
  cols_label(
    y.level = "Outcome",
    term = "Term",
    estimate = "Estimate",
    conf.low = "CI Lower",
    conf.high = "CI Upper")
```

## Prediction about voting preference
```{r}
plot_predictions(fit_nes, 
                 by = "sex",
                 type = "prob",
                 draw = FALSE) |>
# Create the plot
ggplot(aes(x = group, y = estimate, fill = sex)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                position = position_dodge(width = 0.8), width = 0.2) +
  labs(
    title = "Voting Preference by Candidate and Sex",
    subtitle = "Female respondents were more likely to support Clinton; males leaned more toward Perot relative to Bush",
    caption = "Source: American National Election Studies: 2022",
    x = "Candidate",
    y = "Estimated Probability of Vote") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_manual(values = c("Female" = "#D55E00", "Male" = "#0072B2")) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 12, margin = margin(b = 10)),
    plot.caption = element_text(size = 10, face = "italic"),
    axis.title = element_text(face = "bold"),
    legend.title = element_blank())                    
```

## Description of the Chart
<div style="text-align: justify;">The chart illustrates the estimated probability of voting in 1992 Presidential elections for three candidates—Bush, Clinton, and Perot—based on respondents’ sex, using data from the American National Election Studies (2022). Each candidate is represented on the x-axis with two adjacent bars: orange for Female respondents and blue for Male respondents. The y-axis shows the estimated probability of vote, ranging from 0% to over 60%, and includes error bars to reflect uncertainty or variability in the estimates.

The chart reveals clear gender-based differences in voting preferences. Female respondents showed a significantly higher estimated probability of voting for Clinton compared to males, with Clinton receiving the highest overall support among females—exceeding 50%. In contrast, male respondents were more inclined to support Perot than females, whose support for him was notably lower. Support for Bush was relatively similar between the sexes, with a slight edge for males
</div>